---
title: "START Xchrom analysis"
subtitle: "what we found this far"
author: "Julia Romanowska"
date: 'January 7, 2022'
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r knitr_setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  dpi = 150
)
```

```{r data_setup, include=FALSE}
library(fontawesome)
library(tidyverse)
library(patchwork)
# library(karyoploteR) # <- I'm using their functions, but explicitly since
                       # there was some crash of these functions with another pkg
library(sanzo)
library(ggrepel)
library(UpSetR)

signif_pval_thresh <- 0.05
dmr_signif_pval_thresh <- 0.01
dmr_min_n_cpgs <- 3

data_folder <- file.path("/home", "jro049", "Naukowe", "START_project",
                         "X-chrom", "DATA")
source(file.path("/home", "jro049", "Naukowe", "START_project",
                         "X-chrom", "gg_qqplot.R"))

# here are the model names as they appear in the raw data,
#  the order signifies the number of model in the final data presentation
all_model_names <- c("1c", "parents.1c", "1d.bw", "parents.1d.bw")
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
# setting up extra features:
library(xaringanExtra)
use_scribble()
use_tachyons()
use_fit_screen()

library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

<!-- - why are there many more DMRs found in girls-only analyses? -->

**Who wants to be on the author list?**

<div style="float: right; display: inline; font-size: small;">

<p style="font-style: bold;"> LEGEND:</p>

<table style="width: 250px; right: 50px;">
  <tr>
    <th>symbol</th>
    <th style="border-left: 1px solid black;">responsibilities</th>
  </tr>
  <tr>
    <td style="text-align: center;">`r fa("child")`</td><td>playing with data</td>
  </tr>
  <tr>
    <td style="text-align: center;">`r fa("user-md")`</td><td>contribute to discussion; biological expertise</td>
  </tr>
  <tr>
    <td style="text-align: center;">`r fa("user-secret")`</td><td>checking literature, making text shine</td>
  </tr>
  <tr>
    <td style="text-align: center;">`r fa("user-graduate")`</td><td>supervision</td>
  </tr>
  <tr>
    <td style="text-align: center;">`r fa("user-astronaut")`</td><td>suddenly appearing and asking <i>the most difficult question</i></td>
  </tr>
  <tr>
    <td style="text-align: center;">`r fa("user-friends")`</td><td>friendly reviewing the bioinf. approach and implementation</td>
  </tr>
</table>

</div>

<br>
currently:

- `r fa("child")` *Julia Romanowska*,

- `r fa("child")` *Haakon E. Nustad*,

- `r fa("user-friends")` *Christian M. Page*,

- `r fa("user-friends")` *William R.P. Denault*,

- `r fa("user-friends")` *Jon Bohlin*,

- `r fa("user-friends")` *Yunsung Lee*,

- `r fa("user-md")` *Maria C. Magnus*,

- `r fa("user-astronaut")` *Håkon K. Gjessing*,

- `r fa("user-md")` *Robert Lyle*,

- `r fa("user-graduate")` *Per Magnus*,

- `r fa("user-graduate")` *Siri E. Håberg*,

- `r fa("user-secret")` *Astanand Jugessur*

