---
title: "RnBeads pipeline example"
author: "Julia Romanowska"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set( echo = TRUE )
```

# Preparations

Load the libraries and set necessary variables, file paths, etc.

```{r libraries,eval=FALSE}
if ( !requireNamespace( "BiocManager", quietly = TRUE ) ){
    install.packages("BiocManager")
}
BiocManager::install( "RnBeads", version = "3.8" )
BiocManager::install( "RnBeads.hg19", version = "3.8" )
```


```{r setup_paths}
suppressMessages( require( RnBeads ) )

this.dir <- getwd()
idat.dir <- file.path( this.dir, "TEST_DATA" )
sample.annotation <- file.path( this.dir, "TEST_DATA", "sample_sheet.csv" )
analysis.dir <- file.path( this.dir, "RnBeads_analysis" )
report.dir <- file.path( this.dir, "RnBeads_reports2" )
```

Important parameters of the run (to see all the parameters, run `?rnb.options`):

```{r setup_args}
rnb.options(
#  disk.dump.big.matrices = TRUE, ## to save memory
#  disk.dump.bigff = TRUE, ## to save memory
#  logging.disk = TRUE, ## to save memory
#  enforce.memory.management = TRUE, ## to save memory
  analysis.name = "CordBlood",
  identifiers.column = "Sample_Name",
  normalization.method = "bmiq",  ## To be discussed, but probably the best normalizing for now
#  filtering.blacklist = "48639-non-specific-probes-Illumina450k.txt", ## CMP will provide the list 
  filtering.greedycut.pvalue.threshold = 0.01, ## Specific value is probably ok, but should be discussed
  filtering.sex.chromosomes.removal = FALSE,
  exploratory = TRUE,
  exploratory.columns = c( "Sentrix_ID","Sentrix_Position","Year","Age" ),
  exploratory.beta.distribution = FALSE,
  exploratory.clustering = "none",
  min.group.size = 1, ## this is to allow for 'Age' column in exploratory analysis
  differential = FALSE
)
```

Why not "filtering.cross.reactive = TRUE"?

## Correcting the sample annotation file

There was an error when I ran the `rnb.run.analysis`, which stated:

```
2019-02-18 14:35:07     1.6  STATUS STARTED RnBeads Pipeline
2019-02-18 14:35:07     1.6    INFO     Analysis Title: CordBlood
2019-02-18 14:35:07     1.6    INFO     Initialized report index and saved to index.html
2019-02-18 14:35:07     1.6  STATUS     STARTED Loading Data
2019-02-18 14:35:07     1.6    INFO         Number of cores: 1
2019-02-18 14:35:07     1.6    INFO         Loading data of type "idat.dir"
2019-02-18 14:35:07     1.6  STATUS         STARTED Loading Data from IDAT Files
2019-02-18 14:35:07     1.6    INFO             Added column barcode to the provided sample annotation table
2019-02-18 14:35:07     1.6   ERROR             Some IDAT files are not present in the supplied base directory, for instance 201236480142_R01C01_Red.idat, 201236480142_R01C01_Grn.idat, 201236480142_R02C01_Red.idat, 201236480142_R02C01_Grn.idat, 201236480142_R03C01_Red.idat, 201236480142_R03C01_Grn.idat
Error in logger.error(txt) : 
  Some IDAT files are not present in the supplied base directory, for instance 201236480142_R01C01_Red.idat, 201236480142_R01C01_Grn.idat, 201236480142_R02C01_Red.idat, 201236480142_R02C01_Grn.idat, 201236480142_R03C01_Red.idat, 201236480142_R03C01_Grn.idat
```

I need to check which files are missing:

+ load the `sample_sheet.csv` file:

```{r}
sample.sheet <- read.csv( "TEST_DATA/sample_sheet.csv", stringsAsFactors = FALSE )
head( sample.sheet )
head( list.files( "TEST_DATA/" ) )
```

The filenames are constructed as follows:

  - sample name
  - sentrix ID
  - sentrix position
  - green or red channel

+ need to check which files are missing, based on the list in `sample.sheet`:

```{r,eval=FALSE}
all.possible.filenames <- with( sample.sheet,
  paste0( rep( paste( Sample_Name, Sentrix_ID, Sentrix_Position, sep = "_" ), 2 ),
    c( "_Grn.idat", "_Red.idat" )
  )
)
existing.filenames <- list.files( "TEST_DATA/" )[ grep( ".idat", list.files( "TEST_DATA/" ), fixed = TRUE ) ]

missing.filenames <- all.possible.filenames[ is.na( match( all.possible.filenames, existing.filenames ) ) ]
```

No files are missing, but the naming scheme is not the one used by default in RnBeads! Need to remove the `sample name` bit:

```{r,eval=FALSE}
new.filenames <- with( sample.sheet,
  paste0( rep( paste( Sentrix_ID, Sentrix_Position, sep = "_" ), 2 ),
    c( "_Grn.idat", "_Red.idat" )
  )
)
setwd( "TEST_DATA/" )
invisible( mapply( FUN = function( old.fn, new.fn ){
  system( paste( "mv", old.fn, new.fn ) )
}, existing.filenames, new.filenames ) )
setwd( "../" )
```

# Running the analysis

If you run on a multi-core computer with a lot of memory, you can enable parallel mode:
```{r parallel_setup}
logger.start( fname = NA )
num.cores <- 8
parallel.setup( num.cores )
parallel.isEnabled()
```

```{r run}
rnbset.analysed <- rnb.run.analysis(
    dir.reports = report.dir, 
    sample.sheet= sample.annotation, 
    data.dir    = idat.dir, 
    data.type   = "idat.dir" )

parallel.teardown()
```

